---
title: "Seminar 1"
author: "Shiva Chelluri"
date: "`r Sys.Date()`"
output: pdf_document
---

# Introduction:

You can conduct basic math as follows:

```{r}
# you can conduct basic math
x <- 3 + sin(pi/2)

# you can also conduct more complicated mathematical 
sqrt(x)

# you can also save strings dynamically
z <- "word"
```

## Sequences:

Sequences are functional in programming such as looping. They can be made using the `seq()` function.

```{r}
v <- seq(1, 10)
# another easier way of making sequences:
v1 <- 1:10
```

## Vectors:

You can create vectors of data, which can include numbers, characters and other kinds of types as well:

```{r}
w <- c(12, 1, 2, 3)

# you can also sort vectors using the sort() function
sort(w)

```

As you can see, `w` itself does not change because the `sort()` function does apply in place. So, to store this newer version, you need to save it again:

```{r}
w <- sort(w)
```

You can also do **indexing over a vector:**

```{r}
w[1]
```

Note that the indexing in R starts with one and not zero, like in other languages. Additionally, you can conduct scalar multiplications on vectors:

```{r}
w*2
```

You conduct more complex matrix methods such as inner product and matrix multiplications.

```{r}
# for doing inner-products
w %*% w
```

## Data Formatting:

Usually, you will get data with the wrong data types, and you need to format the data before such as:

```{r}
A <- as.character(c(1, 2, 3, 4, 5))
typeof(A)
```

Here, you have vector numbers, but they are saved as characters. So, we use two auxiliary functions to identify two main attributes of a variable:

1.  Length of the vector: `length()` returns the length of the vector.
2.  Data Type of the vector elements: `typeof()` returns the variable's data type.

```{r}
length_a <- length(A)
typeof_a <- typeof(A)
print(paste0("The vector A is of length", length_a, "and has the data type of ", typeof(A)))
```

**Coercion:** It is the process of forcing data of one datatype to another datatype

### Loops:

You can also conduct for-loops using the `for(){}` function to write loops. They have the common syntax as:

```{r}
for (i in 1:10){
  print(paste0("Printing loop:", i))
}
```

## Packages:

To get external functions, we need to use the `install.packages("")` function to install the packages.

We will primarily use the `tidyverse()` package because it contains most of the other packages for the stuff we are going to do.

```{r}
# loading the tidyverse package
library(tidyverse)
```

### Plotting with `ggplot2()`

Let us load the `mpg` data set from the `tidyverse` package. And let us conduct basic plotting:

```{r}
cars <- mpg


# example ggplot with linear regressions
ggplot(data = cars) +
  geom_point(mapping = aes(x=displ, y=hwy, xlab= "Display", ylab="Highway"))+
  geom_smooth(method = "lm", mapping = aes(x=displ, y=hwy))+
  labs(y="Highway", x = "display", title="My first Ggplot2")+
  theme_minimal()
```

## Introduction to Markdown:

You can write bulleted lists as:

-   Item 1

-   item 2

-   item 2

You can also write numbered lists as:

1.  Item 1
2.  Item 2
3.  Item 3

### Code Blocks:

You can write code-blocks using the ```` ```{r} ```` and it looks like the following:

```{r}
print("This is a code block")
```

### Math:

You can write two types of math:

1.  Inline math: You do it using "\\\$\\\$" symbols. It works such as $y = mx + c$

2.  MathJAX Block: You can do it using double dollar signs:

    $$
    \textbf{Y} = \textbf{X}\beta + \epsilon
    $$\
